A SAT-Based Counterexample Guided Method for Unbounded Synthesis

### 2016-03-24 paper.tex
Added k as an input to height

An abstraction of the game restricts actions available to one of the players.
Specifically, we consider abstractions represented as trees of actions,
referred to as \emph{abstract game trees} (AGTs).  Figure~\ref{fig:agt} shows an
example abstract game tree for the controller (abstract game trees for the
environment are similar).  In the abstract game, the controller is required to
pick actions from the tree, starting from the root node.  After reaching a
leaf, it continues playing unrestricted.  The tree in Figure~\ref{fig:agt}
restricts the initial controller action to the set $\{a, d\}$.  After choosing
action $a$, the controller is required to play action $b$ in the next round.
It then reaches a leaf of the tree and continues playing unrestricted.  The
root of the tree is annotated by the initial state $s$ of the abstract game and
the bound $k$ on the number of rounds.  We denote $\textsc{nodes}(T)$ the set
of all nodes of a tree $T$, $\textsc{leaves}(T)$ the subset of leaf nodes.  For
edge $e$, $\textsc{action}(e)$ is the action that labels the edge, and for node
$n$, $\textsc{height}(n)$ is the maximal distance from n to a leaf. $\textsc{height}(T)$ is the height 
of the root node of the tree.  For node
$n$ of the tree, $\textsc{succ}(n)$ is the set of pairs $\langle e, n' \rangle$
where $n'$ is a child node of $n$ and $e$ is the edge connecting $n$ and $n'$.

@@

    \begin{algorithmic}
        \algrestore{b1}
        \Function{findCandidate}{$p, s, k, T$}
        \State $\hat{T} \gets $ \Call{extend}{$T$} \Comment{Extend the tree with unfixed actions}
            \State $f \gets $ \IfElse{$p = \texttt{cont}$}{\Call{treeFormula}{$\hat{T}$}}{\Call{\textoverline{treeFormula}}{$\hat{T}$}} \EndIfElse
            \State $sol \gets $ \Call{SAT}{$s(X_{\hat{T}}) \land f$}
            \If{$sol = \texttt{unsat}$} 
                \If{\texttt{unbounded}} \Comment{Active only in the unbounded solver}
                    %\State $\sigma \gets $ \Call{generalise}{$s$} \Comment{Expand $s$ to a set of states}
                    \State \IfElse{$p = \texttt{cont}$}{\Call{learn}{$s, \hat{T}$}}{\Call{\textoverline{learn}}{$s, \hat{T}$}} \EndIfElse
                \EndIf
                \State \Return $\texttt{NULL}$ \Comment{No candidate exists}
            \Else
                \State \Return $\{ \langle n, c \rangle | n \in $ \Call{nodes}{$T$} $, c = \Call{sol}{n} \}$ \Comment{Fix candidate moves in $T$}
            \EndIf
        \EndFunction
        \algstore{b2}
    \end{algorithmic}

@@

\begin{algorithm}
    \caption{Tree formulas for Controller and Environment respectively}
    \label{alg:treeFormula}
    \begin{algorithmic}[1]
        \Function{treeFormula}{$T$}
        \If{$\Call{height}{T} = 0$}
        \State \Return{ $\lnot \Call{E}{X_{T}}$ }
        \Else
        \State \Return{$\lnot \Call{E}{X_{T}} \land$ \\
            $$\bigwedge_{\langle e, n \rangle \in \Call{succ}{T}}(\Call{$\delta$}{X_T, U_e, C_T, X_n} \land U_e = \Call{action}{e} \land \Call{treeFormula}{n})$$
        }
        \EndIf
        \EndFunction
        \algstore{tf1}
    \end{algorithmic}

    \begin{algorithmic}[1]
        \algrestore{tf1}
        \Function{\textoverline{treeFormula}}{$T$}
        \If{$\Call{height}{T} = 0$}
        \State \Return{\Call{E}{$X_{T}$}}
        \Else
        \State \Return{ $\Call{E}{X_{T}} \lor$ \\
        $$\bigvee_{\langle e, n \rangle \in \Call{succ}{T}}(\Call{$\delta$}{X_T, U_T, C_e, X_n} \land C_e = \Call{action}{e} \land \Call{\textoverline{treeFormula}}{n})$$ }
        \EndIf
        \EndFunction
    \end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}[t]
    \begin{algorithmic}[1]
        \Require $s(X_T) \land $ \Call{treeFormula}{$T$} $\equiv \bot$
        \Require \emph{Must-invariant} holds
        \Ensure \emph{Must-invariant} holds
        \Ensure $s(X_T) \land B^M \not\equiv \bot$
        \Comment $s$ will be added to $B^M$
        \Function{learn}{$s, T$}
            \IIf{$T = \emptyset$}
                \Return
            \EndIIf
            \State $n \gets $ non-leaf node with min height 
            \State $\langle T_1, T_2 \rangle \gets $ \Call{gtSplit}{$T, n$}
            \State $\mathcal{I} \gets $ \Call{interpolate}{$s(X_T) \land $ \Call{treeFormula}{$T_1$}, \Call{treeFormula}{$T_2$}}
            \State $B^M \gets B^M \lor \mathcal{I}$
            \State \Call{learn}{$s, T_1$}
        \EndFunction
        \algstore{learn}
    \end{algorithmic}

    \begin{algorithmic}[1]
        \algrestore{learn}
        \Require $s(X_T) \land $ \Call{\textoverline{treeFormula}}{$T$} $\equiv \bot$
        \Require \emph{May-invariant} holds
        \Ensure \emph{May-invariant} holds
        \Ensure $s(X_T) \land B^m[$\Call{height}{$T$}$] \equiv \bot$
        \Comment{$s$ will be removed from $B^m[$\Call{height}{$T$}$]$}
        \Function{\textoverline{learn}}{$s, T$}
            \IIf{$T = \emptyset$}
                \Return
            \EndIIf
            \State $n \gets $ non-leaf node with min height 
            \State $\langle T_1, T_2 \rangle \gets $ \Call{gtSplit}{$T, n$}
            \State $\mathcal{I} \gets $ \Call{interpolate}{$s(X_T) \land $ \Call{\textoverline{treeFormula}}{$T_1$}, \Call{\textoverline{treeFormula}}{$T_2$}}
            \For{$i = 1$ to \Call{height}{$n$}}
                \State $B^m[i] \gets B^m[i] \setminus \mathcal{I}$
            \EndFor
            \State \Call{\textoverline{learn}}{$s, T_1$}
        \EndFunction
    \end{algorithmic}
    \caption{Learning algorithms}
    \label{alg:learn}
\end{algorithm}

Consider the game tree $T$ in
Figure~\ref{fig:interpolatetree}, whose edges are labelled with environment
actions.  Assume that there does not exist a partial strategy for the controller
winning against this tree and that the \textsc{findCandidate} function could
not find a candidate strategy. We choose a non-leaf node $n$ of $T$ with
maximal depth, i.e., a node whose children are leafs
(Algorithm~\ref{alg:learn}, line~3). We then split the tree at $n$ such that
both slices $T_1$ and $T_2$ contain a copy of $n$ (line~4).
Figure~\ref{fig:treef1} shows $T_1$, which contains all of $T$ except $n$'s
children, and $T_2$ (Figure~\ref{fig:treef2}), which contains only $n$ and its
children.  There is no candidate strategy for $T$ so $s \land
\textsc{treeFormula}(T)$ is unsatisfiable.  By construction,
$\textsc{treeFormula}(T) \equiv \textsc{treeFormula}(T_1) \land
\textsc{treeFormula}(T_2)$ and hence $s \land \textsc{treeFormula}(T_1) \land
\textsc{treeFormula}(T_2)$ is also unsatisfiable. This enables the construction
of an \emph{interpolant} that captures losing states at node $n$.

@@

We construct an interpolant with $F_1 = s(X_T) \land \textsc{treeFormula}(T_1)$
and $F_2 = \textsc{treeFormula}(T_2)$ (line~5). The only variables shared between $F_1$
and $F_2$ are the state variable copies belonging to node $n$. By the properties of the interpolant,
$F_2 \land \mathcal{I}$ is unsatisfiable, therefore all states in $\mathcal{I}$
are losing against abstract game tree $T_2$ in Figure~\ref{fig:treef2}.  We also know that $F_1 \to
\mathcal{I}$, thus $\mathcal{I}$ contains all states reachable at $n$ by
following $T_1$ and avoiding error states.

@@

\begin{algorithm}[t]
    \caption{Amended tree formulas for Controller and Environment}
    \label{alg:unboundedTreeFormula}
    \begin{algorithmic}[1]
        \Function{treeFormula}{$T$}
        \If{$\Call{height}{T} = 0$}
        \State \Return{ $\lnot \Call{$B^M$}{X_{T}}$ }
        \Else
        \State \Return{$\lnot \Call{$B^M$}{X_{T}} \land$ \\
            $$\bigwedge_{\langle e, n \rangle \in \Call{succ}{T}}(\Call{$\delta$}{X_T, U_e, C_n, X_n} \land U_e = \Call{action}{e} \land \Call{treeFormula}{n})$$
        }
        \EndIf
        \EndFunction
        \algstore{tf1}
    \end{algorithmic}

    \begin{algorithmic}[1]
        \algrestore{tf1}
        \Function{\textoverline{treeFormula}}{$T$}
        \If{$\Call{height}{T} = 0$}
        \State \Return{\Call{E}{$X_T$}}
        \Else
        \State \Return{ \Call{$B^m[\Call{height}{T}]$}{$X_{T}$} $\land$ \\
            $$\bigg( \Call{E}{X_T} \lor \bigvee_{\langle e, n \rangle \in \Call{succ}{T}}(\Call{$\delta$}{X_T, U_n, C_e, X_n} \land C_e = \Call{action}{e} \land \Call{\textoverline{treeFormula}}{n})\bigg)$$ }
        \EndIf
        \EndFunction
    \end{algorithmic}
\end{algorithm}

Learning of states losing for the environment is similar (\textsc{\textoverline{learn}} in Algorithm~\ref{alg:learn}). 
The main difference is that environment-losing states are only losing for a particular
bound: given that there does not exist an environment strategy that
forces the game into an error state in $k$ rounds or less; there may still exist a longer
environment-winning strategy.  We therefore record learned environment-losing states
along with associated bounds.  To this end, we maintain a conceptually infinite array of
sets $B^m[k]$ that are may-losing for the controller, indexed by bound $k$.  $B^m[k]$ 
are initialised to $\top$ for all $k$.  Whenever an environment-losing set $\mathcal{I}$
is discovered for a node $n$ with bound $\textsc{height}(n)$ in line~13 of Algorithm~\ref{alg:learn}, this set is subtracted from 
$B^m[i]$, for all $i$ less than or equal to the bound (lines~14--16).

@@

Line~(11--12) splits the tree $T$ into $T_1$ and $T_2$, such that $T_2$ has depth
1.  Consider formulas $F_1=s(X_T) \land
\textsc{\textoverline{treeFormula}}(T_1)$ and $F_2 =
\textsc{\textoverline{treeFormula}}(T_2)$.  These formulas only share variables
$X_n$.  Their conjunction $F_1 \land F_2$ is unsatisfiable, as by construction
any solution of $F_1 \land F_2$ also satisfies $s(X_T) \land
\textsc{\textoverline{treeFormula}}(T)$, which is unsatisfiable (precondition (b)).  Hence the
interpolation operation is defined for $F_1$ and $F_2$.

@@

Formally, $\II$ has the property $\II \land F_2 \equiv \bot$.  Since $T_2$ is
of depth 1, this means that the environment cannot force the game into
$B^m[\textsc{height}(n)-1]$ playing against the counterexample moves in $T_2$.
Hence, $\II \cap Upre(B^m[\textsc{height}(n)-1]) = \emptyset$.  Furthermore,
since the may-invariant holds, $\II \cap Upre(B^m[\textsc{height}(i)]) =
\emptyset$, for all $i < \textsc{height}(n)$.  Hence, removing $\II$ from all
$B^m[i], i\leq \textsc{height}(n)$ in line~(15) preserves the may-invariant,
thus satisfying the first post-condition.

Furthermore, the interpolant satisfies $F_1 \rightarrow \II$, i.e., any
assignment to $X_n$ that satisfies $s(X_T) \land
\textsc{\textoverline{treeFormula}}(T_1)$ also satisfies $\II$.  Hence,
removing $\II$ from $B^m[\textsc{height}(n)]$ makes $s(X_T) \land
\textsc{\textoverline{treeFormula}}(T_1)$ unsatisfiable, and hence all
preconditions of the recursive invocation of \textsc{\textoverline{learn}} in
line~(17) are satisfied.

At the second last recursive call to \textsc{\textoverline{learn}}, tree $T_1$
is empty, $n$ is the root node, $\textsc{\textoverline{treeFormula}}(T_1)
\equiv B^m[\textsc{height}(T_1)](X^T)$; hence $s(X_T) \land
\textsc{\textoverline{treeFormula}}(T_1) \equiv s(X_T) \land
B^m[\textsc{height}(T_1)](X^T) \equiv \bot$.  Thus the second postcondition of
\textsc{\textoverline{learn}} holds.

The proof of \textsc{learn} is similar to the above proof of \textsc{learn}. An
interpolant constructed from $F_1=s(X_T) \land \textsc{treeFormula}(T_1)$
and $F_2 = \textsc{treeFormula}(T_2)$ has the property $\II \land F_2 \equiv
\bot$ and the precondition ensures that the controller is unable to force the
game into $B^M$ playing against the counterexample moves in $T_2$. Thus adding
$\II$ to $B^M$ maintains the must-invariant satisfying the first postcondition.

Likewise, in the second last recursive call of \textsc{learn} with the empty
tree $T_1$ and root node $n$: $\textsc{treeFormula}(T_1) \equiv \lnot
B^M(X_T)$.  Hence $s(X_T) \land \textsc{treeFormula}(T_1) \equiv
s(X_T) \land \lnot B^M(X_T) \equiv \bot$. Therefore $s(X_T) \land
B^M(X_T) \not\equiv \bot$, the second postcondition, is true.

>>>>

An abstraction of the game restricts actions available to one of the players.
Specifically, we consider abstractions represented as trees of actions,
referred to as \emph{abstract game trees} (AGTs).  Figure~\ref{fig:agt} shows an
example abstract game tree for the controller (abstract game trees for the
environment are similar).  In the abstract game, the controller is required to
pick actions from the tree, starting from the root node.  After reaching a
leaf, it continues playing unrestricted.  The tree in Figure~\ref{fig:agt}
restricts the initial controller action to the set $\{a, d\}$.  After choosing
action $a$, the controller is required to play action $b$ in the next round.
It then reaches a leaf of the tree and continues playing unrestricted.  The
root of the tree is annotated by the initial state $s$ of the abstract game and
the bound $k$ on the number of rounds.  We denote $\textsc{nodes}(T)$ the set
of all nodes of a tree $T$, $\textsc{leaves}(T)$ the subset of leaf nodes.  For
edge $e$, $\textsc{action}(e)$ is the action that labels the edge, and for node
$n$, $\textsc{height}(k, n)$ is the distance from n to the last round of a game
bounded to $k$ rounds.  $\textsc{height}(k, T)$ is the height of the root node
of the tree.  For node $n$ of the tree, $\textsc{succ}(n)$ is the set of pairs
$\langle e, n' \rangle$ where $n'$ is a child node of $n$ and $e$ is the edge
connecting $n$ and $n'$.

@@

    \begin{algorithmic}
        \algrestore{b1}
        \Function{findCandidate}{$p, s, k, T$}
        \State $\hat{T} \gets $ \Call{extend}{$T$} \Comment{Extend the tree with unfixed actions}
            \State $f \gets $ \IfElse{$p = \texttt{cont}$}{\Call{treeFormula}{$k, \hat{T}$}}{\Call{\textoverline{treeFormula}}{$k, \hat{T}$}} \EndIfElse
            \State $sol \gets $ \Call{SAT}{$s(X_{\hat{T}}) \land f$}
            \If{$sol = \texttt{unsat}$} 
                \If{\texttt{unbounded}} \Comment{Active only in the unbounded solver}
                    %\State $\sigma \gets $ \Call{generalise}{$s$} \Comment{Expand $s$ to a set of states}
                    \State \IfElse{$p = \texttt{cont}$}{\Call{learn}{$s, \hat{T}$}}{\Call{\textoverline{learn}}{$s, \hat{T}$}} \EndIfElse
                \EndIf
                \State \Return $\texttt{NULL}$ \Comment{No candidate exists}
            \Else
                \State \Return $\{ \langle n, c \rangle | n \in $ \Call{nodes}{$T$} $, c = \Call{sol}{n} \}$ \Comment{Fix candidate moves in $T$}
            \EndIf
        \EndFunction
        \algstore{b2}
    \end{algorithmic}

@@

\begin{algorithm}
    \caption{Tree formulas for Controller and Environment respectively}
    \label{alg:treeFormula}
    \begin{algorithmic}[1]
        \Function{treeFormula}{$k, T$}
        \If{$\Call{height}{k, T} = 0$}
        \State \Return{ $\lnot \Call{E}{X_{T}}$ }
        \Else
        \State \Return{$\lnot \Call{E}{X_{T}} \land$ \\
            $$\bigwedge_{\langle e, n \rangle \in \Call{succ}{T}}(\Call{$\delta$}{X_T, U_e, C_T, X_n} \land U_e = \Call{action}{e} \land \Call{treeFormula}{k, n})$$
        }
        \EndIf
        \EndFunction
        \algstore{tf1}
    \end{algorithmic}

    \begin{algorithmic}[1]
        \algrestore{tf1}
        \Function{\textoverline{treeFormula}}{$k, T$}
        \If{$\Call{height}{k, T} = 0$}
        \State \Return{\Call{E}{$X_{T}$}}
        \Else
        \State \Return{ $\Call{E}{X_{T}} \lor$ \\
        $$\bigvee_{\langle e, n \rangle \in \Call{succ}{T}}(\Call{$\delta$}{X_T, U_T, C_e, X_n} \land C_e = \Call{action}{e} \land \Call{\textoverline{treeFormula}}{k, n})$$ }
        \EndIf
        \EndFunction
    \end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}[t]
    \begin{algorithmic}[1]
        \Require $s(X_T) \land $ \Call{treeFormula}{$k, T$} $\equiv \bot$
        \Require \emph{Must-invariant} holds
        \Ensure \emph{Must-invariant} holds
        \Ensure $s(X_T) \land B^M \not\equiv \bot$
        \Comment $s$ will be added to $B^M$
        \Function{learn}{$s, T$}
            \IIf{$T = \emptyset$}
                \Return
            \EndIIf
            \State $n \gets $ non-leaf node with min height 
            \State $\langle T_1, T_2 \rangle \gets $ \Call{gtSplit}{$T, n$}
            \State $\mathcal{I} \gets $ \Call{interpolate}{$s(X_T) \land $ \Call{treeFormula}{$k, T_1$}, \Call{treeFormula}{$k, T_2$}}
            \State $B^M \gets B^M \lor \mathcal{I}$
            \State \Call{learn}{$s, T_1$}
        \EndFunction
        \algstore{learn}
    \end{algorithmic}

    \begin{algorithmic}[1]
        \algrestore{learn}
        \Require $s(X_T) \land $ \Call{\textoverline{treeFormula}}{$k, T$} $\equiv \bot$
        \Require \emph{May-invariant} holds
        \Ensure \emph{May-invariant} holds
        \Ensure $s(X_T) \land B^m[$\Call{height}{$k, T$}$] \equiv \bot$
        \Comment{$s$ will be removed from $B^m[$\Call{height}{$k, T$}$]$}
        \Function{\textoverline{learn}}{$s, T$}
            \IIf{$T = \emptyset$}
                \Return
            \EndIIf
            \State $n \gets $ non-leaf node with min height 
            \State $\langle T_1, T_2 \rangle \gets $ \Call{gtSplit}{$T, n$}
            \State $\mathcal{I} \gets $ \Call{interpolate}{$s(X_T) \land $ \Call{\textoverline{treeFormula}}{$k, T_1$}, \Call{\textoverline{treeFormula}}{$k, T_2$}}
            \For{$i = 1$ to \Call{height}{$k, n$}}
                \State $B^m[i] \gets B^m[i] \setminus \mathcal{I}$
            \EndFor
            \State \Call{\textoverline{learn}}{$s, T_1$}
        \EndFunction
    \end{algorithmic}
    \caption{Learning algorithms}
    \label{alg:learn}
\end{algorithm}

Consider the game tree $T$ in
Figure~\ref{fig:interpolatetree}, whose edges are labelled with environment
actions.  Assume that there does not exist a partial strategy for the controller
winning against this tree and that the \textsc{findCandidate} function could
not find a candidate strategy. We choose a non-leaf node $n$ of $T$ with
maximal depth, i.e., a node whose children are leafs
(Algorithm~\ref{alg:learn}, line~3). We then split the tree at $n$ such that
both slices $T_1$ and $T_2$ contain a copy of $n$ (line~4).
Figure~\ref{fig:treef1} shows $T_1$, which contains all of $T$ except $n$'s
children, and $T_2$ (Figure~\ref{fig:treef2}), which contains only $n$ and its
children.  There is no candidate strategy for $T$ so $s \land
\textsc{treeFormula}(k, T)$ is unsatisfiable.  By construction,
$\textsc{treeFormula}(k, T) \equiv \textsc{treeFormula}(k, T_1) \land
\textsc{treeFormula}(k, T_2)$ and hence $s \land \textsc{treeFormula}(k, T_1) \land
\textsc{treeFormula}(k, T_2)$ is also unsatisfiable. This enables the construction
of an \emph{interpolant} that captures losing states at node $n$.

@@

We construct an interpolant with $F_1 = s(X_T) \land \textsc{treeFormula}(k, T_1)$
and $F_2 = \textsc{treeFormula}(k, T_2)$ (line~5). The only variables shared between $F_1$
and $F_2$ are the state variable copies belonging to node $n$. By the properties of the interpolant,
$F_2 \land \mathcal{I}$ is unsatisfiable, therefore all states in $\mathcal{I}$
are losing against abstract game tree $T_2$ in Figure~\ref{fig:treef2}.  We also know that $F_1 \to
\mathcal{I}$, thus $\mathcal{I}$ contains all states reachable at $n$ by
following $T_1$ and avoiding error states.

@@

\begin{algorithm}[t]
    \caption{Amended tree formulas for Controller and Environment}
    \label{alg:unboundedTreeFormula}
    \begin{algorithmic}[1]
        \Function{treeFormula}{$k, T$}
        \If{$\Call{height}{k, T} = 0$}
        \State \Return{ $\lnot \Call{$B^M$}{X_{T}}$ }
        \Else
        \State \Return{$\lnot \Call{$B^M$}{X_{T}} \land$ \\
            $$\bigwedge_{\langle e, n \rangle \in \Call{succ}{T}}(\Call{$\delta$}{X_T, U_e, C_n, X_n} \land U_e = \Call{action}{e} \land \Call{treeFormula}{k, n})$$
        }
        \EndIf
        \EndFunction
        \algstore{tf1}
    \end{algorithmic}

    \begin{algorithmic}[1]
        \algrestore{tf1}
        \Function{\textoverline{treeFormula}}{$k, T$}
        \If{$\Call{height}{k, T} = 0$}
        \State \Return{\Call{E}{$X_T$}}
        \Else
        \State \Return{ \Call{$B^m[\Call{height}{k, T}]$}{$X_{T}$} $\land$ \\
            $$\bigg( \Call{E}{X_T} \lor \bigvee_{\langle e, n \rangle \in \Call{succ}{T}}(\Call{$\delta$}{X_T, U_n, C_e, X_n} \land C_e = \Call{action}{e} \land \Call{\textoverline{treeFormula}}{k, n})\bigg)$$ }
        \EndIf
        \EndFunction
    \end{algorithmic}
\end{algorithm}

Learning of states losing for the environment is similar (\textsc{\textoverline{learn}} in Algorithm~\ref{alg:learn}). 
The main difference is that environment-losing states are only losing for a particular
bound: given that there does not exist an environment strategy that
forces the game into an error state in $k$ rounds or less; there may still exist a longer
environment-winning strategy.  We therefore record learned environment-losing states
along with associated bounds.  To this end, we maintain a conceptually infinite array of
sets $B^m[k]$ that are may-losing for the controller, indexed by bound $k$.  $B^m[k]$ 
are initialised to $\top$ for all $k$.  Whenever an environment-losing set $\mathcal{I}$
is discovered for a node $n$ with bound $\textsc{height}(k, n)$ in line~13 of Algorithm~\ref{alg:learn}, this set is subtracted from 
$B^m[i]$, for all $i$ less than or equal to the bound (lines~14--16).

@@

Line~(11--12) splits the tree $T$ into $T_1$ and $T_2$, such that $T_2$ has depth
1.  Consider formulas $F_1=s(X_T) \land
\textsc{\textoverline{treeFormula}}(k, T_1)$ and $F_2 =
\textsc{\textoverline{treeFormula}}(k, T_2)$.  These formulas only share variables
$X_n$.  Their conjunction $F_1 \land F_2$ is unsatisfiable, as by construction
any solution of $F_1 \land F_2$ also satisfies $s(X_T) \land
\textsc{\textoverline{treeFormula}}(k, T)$, which is unsatisfiable (precondition (b)).  Hence the
interpolation operation is defined for $F_1$ and $F_2$.

@@

Formally, $\II$ has the property $\II \land F_2 \equiv \bot$.  Since $T_2$ is
of depth 1, this means that the environment cannot force the game into
$B^m[\textsc{height}(k, n)-1]$ playing against the counterexample moves in $T_2$.
Hence, $\II \cap Upre(B^m[\textsc{height}(k, n)-1]) = \emptyset$.  Furthermore,
since the may-invariant holds, $\II \cap Upre(B^m[i]) =
\emptyset$, for all $i < \textsc{height}(k, n)$.  Hence, removing $\II$ from all
$B^m[i], i\leq \textsc{height}(k, n)$ in line~(15) preserves the may-invariant,
thus satisfying the first post-condition.

Furthermore, the interpolant satisfies $F_1 \rightarrow \II$, i.e., any
assignment to $X_n$ that satisfies $s(X_T) \land
\textsc{\textoverline{treeFormula}}(k, T_1)$ also satisfies $\II$.  Hence,
removing $\II$ from $B^m[\textsc{height}(k, n)]$ makes $s(X_T) \land
\textsc{\textoverline{treeFormula}}(k, T_1)$ unsatisfiable, and hence all
preconditions of the recursive invocation of \textsc{\textoverline{learn}} in
line~(17) are satisfied.

At the second last recursive call to \textsc{\textoverline{learn}}, tree $T_1$
is empty, $n$ is the root node, $\textsc{\textoverline{treeFormula}}(k, T_1)
\equiv B^m[\textsc{height}(k, T_1)](X^T)$; hence $s(X_T) \land
\textsc{\textoverline{treeFormula}}(k, T_1) \equiv s(X_T) \land
B^m[\textsc{height}(k, T_1)](X^T) \equiv \bot$.  Thus the second postcondition of
\textsc{\textoverline{learn}} holds.

The proof of \textsc{learn} is similar to the above proof of \textsc{learn}. An
interpolant constructed from $F_1=s(X_T) \land \textsc{treeFormula}(k, T_1)$
and $F_2 = \textsc{treeFormula}(k, T_2)$ has the property $\II \land F_2 \equiv
\bot$ and the precondition ensures that the controller is unable to force the
game into $B^M$ playing against the counterexample moves in $T_2$. Thus adding
$\II$ to $B^M$ maintains the must-invariant satisfying the first postcondition.

Likewise, in the second last recursive call of \textsc{learn} with the empty
tree $T_1$ and root node $n$: $\textsc{treeFormula}(k, T_1) \equiv \lnot
B^M(X_T)$.  Hence $s(X_T) \land \textsc{treeFormula}(k, T_1) \equiv
s(X_T) \land \lnot B^M(X_T) \equiv \bot$. Therefore $s(X_T) \land
B^M(X_T) \not\equiv \bot$, the second postcondition, is true.
